Developing effective theories for the square lattice Hubbard model using density matrix downfolding

### 2019-07-26 draft.tex
worked on propsed outline. ready for writing, has TBLG stuff now

\section{Introduction}
A core problem of condensed matter physics is understanding how to systematically develop low-energy effective models for strongly correlated electron systems. 
In the forefront of modern research are materials like high temperature superconductors and heavy fermion systems for which simple effective theories have not yet been found. 
For example, the cuprate family of high-Tc superconductors have an intricate phase diagram in the doping-temperature space with anti-ferromagnetic (AFM) insulating, superconducting, psuedo-gap, "bad metal", and Fermi-liquid like phases of which only the first and last are well understood.
The iron-pnictides also have superconducting domes, non-Fermi liquid like metallic phases, a spin-density-wave antiferommagnetic phase, all under various dopings and temperatures. 
Heavy-fermion materials have a purported quantum critical point between an AFM and Fermi-liquid phase which houses a superconducting dome, leading to another phase diagram for which building model Hamiltonians is very challenging.

@@

\textcolor{red}{
We propose using a density matrix downfolding (DMD) procedure to systematically develop low-energy effective theories of the 1-band nearest-neighbor Hubbard model on a square lattice for various couplings U and temperatures T at half-filling as a stepping stone towards building effective theories for strongly correlated materials. 
The Hubbard model is a simple model with nearest-neighbor hopping and a local on-site Coulomb repulsion, parameterized by a single ratio U/t: 
\begin{equation}
H_\text{hub} = -t \sum_{\langle i,j \rangle,\sigma}( c_{i^\dagger,\sigma} c_{j,\sigma} + h.c.) + U \sum_i n_{i\uparrow} n_{i\downarrow}
\label{hubbard}.
\end{equation}
The solutions in 1-d are well understood via the Bethe ansatz, however, the 2-d square lattice Hubbard model has no known exact solution.  
A few well accepted facts do exist for this model, such as the existence of long-range anti-ferromagnetic (AFM) order for any U$>$0 at T = 0 and a Mott-Heisenberg limit as $U/t \rightarrow \infty$ at low temperatures.
The T = 0 diagram of the model at half-filling has been a long-standing source of controversy between two groups:  those who believe that for any U $>$ 0 the model exhibits Mott-Heisenberg behavior, and those who believe that there is a transition at finite U$^*$ between Slater and Mott-Heisenberg behavior.
While recent numerical and theoretical calculations tend to agree with the latter position, non-negligible evidence for the prior does exist. 
Further, modern finite temperature calculations indicate thermal transitions in the weak to moderate coupling region between the Slater phase at T = 0, pseudogap phase below T$_{pg}$ and a Fermi liquid phase above.
A finite temperature psuedogap to Mott insulator transition has also been found between the weak to strong coupling limits.
The nature of these transitions as phase transitions or simply crossovers is poorly understood.
The relative simplicity of the half-filled Hubbard model, potential complexity of its phase diagram in U-T space,  and the lack of well-established low-energy theories for moderate U/t lends the model as a useful playground for our new downfolding procedure in preparation for tackling realistic strongly correlated electron systems. 
}

\section{Methods}
\subsection{Density Matrix Downfolding (DMD)}
\textbf{We begin by defining what we mean by a low-energy effective Hamiltonian and downfolding.}
Consider a Hamiltonian $\hat{H}$ defined on a Hilbert space $\mathcal{H}$.
A low-energy effective Hamiltonian is an operator $\hat{H}_\text{eff}$ which can accurately reproduce the spectra, up to a constant, and eigenstates of $\hat{H}$ on a subpsace $\mathcal{LE}$ of the Hilbert space $\mathcal{H}$.
Here $\mathcal{LE}$ is taken to be the span of the N lowest energy eigenstates of $\hat{H}$, and $\hat{H}_\text{eff}$ is a linear combination of Hermitian operators $\hat{d}_k$ with a constant energy shift $E_0$
\begin{equation}
\hat{H}_\text{eff} = \sum_{k} g_k \hat{d}_k  + E_0.
\label{eq:Heff}
\end{equation}
The objective of Hamiltonian downfolding is to construct such an $\hat{H}_\text{eff}$ given $\hat{H}, \mathcal{H}$.

\textbf{The key insight of DMD is that the Hamiltonian downfolding problem can be mapped onto a linear regression problem.}
It has been shown that the above definition of $\hat{H}_\text{eff}$ is equivalent to the following: a low-energy effective Hamiltonian is an operator on a subspace $\mathcal{LE}$ of $\mathcal{H}$ such that 
\begin{equation}
\begin{split}
\forall |\Psi\rangle \in \mathcal{LE},\ (E[\Psi] \equiv \langle \Psi|\hat{H} | \Psi \rangle)  = \\ (E_\text{eff}[\Psi] \equiv \langle \Psi | \hat{H}_\text{eff} | \Psi \rangle) + \epsilon[\Psi]
\end{split}
\label{eq:DMD}
\end{equation}
where $\epsilon[\Psi]$ is the error in the effective theory.
Given the general linear form of $\hat{H}_\text{eff}$ seen in \eqref{eq:Heff}, 
the task of building an effective Hamiltonian then reduces to fitting a linear model that minimizes the error $\epsilon[\Psi]$ over $\mathcal{LE}$.
The independent variables (descriptors) in the fitting are expectation values of the operators $d_k[\Psi] \equiv \langle \Psi |\hat{d}_k|\Psi \rangle$ and the target variable is $E[\Psi]$.

\textbf{This linear regression problem can be tackled in three steps, beginning with sampling states from $\mathcal{LE}$.}
Ideally the linear regression would be conducted over the entire space $\mathcal{LE}$. 
In practice, the space is too large, and instead a representative sample set is drawn and used for fitting.
For each sampled state $|\Psi_s\rangle$, the quantities $E[\Psi_s], \{d_k[\Psi_s]\}$ are calculated to be used in the linear regression.
Optimal sampling schemes vary between systems and can significantly reduce the number of samples required to achieve robust statistics.

\textbf{Generally, however, one cannot sample the true low-energy subspace and low-energy wave functions are sampled from an approximate subspace $\mathcal{LE}^\prime$.}
A model fit to an arbitrary subspace of $\mathcal{H}$ will generally fail in describing the low-energy excitations of $\hat{H}$.
Rather, the choice of approximate subspace must be made carefully to ensure the fit model is transferable to the true low-energy space.
We describe in detail later how we selected $\mathcal{LE}^\prime$ for the CuO molecule and the scheme used to draw samples from it.

\textbf{Next, a set of candidate descriptors which form the effective Hamiltonian are selected.}
Descriptors will typically be selected if their values co-vary with the energy over the set of sampled states. 
Knowledge of the low-energy degrees of the freedom from experiment and physical intuition can help restrict the set of candidate descriptors further.
In principle there is no restriction on the form of the candidate operators other than Hermiticity, but second quantized operators are commonly used.

\textbf{Typically, candidate operators are written in second quantized form, and an appropriate single particle basis to express these operators must be constructed.}
An appropriate choice of single particle basis can simplify the representation of the effective theory and typically depends on the system of interest.
Common choices include Bloch waves and molecular orbitals for delocalized basis elements or Wannier and intrinsic atomic orbitals for localized ones.
The selected single particle basis can be validated by ensuring it accurately describes single particle properties among the sampled states like the total number of electrons in the system.

\textbf{ Finally, the sampled data are used to fit the coefficients $\{g_k\}$ by linear regression.}
A typical fitting workflow involves feature selection, model regression and model validation.
Feature selection methods for choosing appropriate descriptors can range from wrapper methods like orthogonal matching pursuit and principal component analysis to embedded methods like LASSO, ridge and elastic net regularizations. 
Fitting the effective Hamiltonian usually involves an ordinary least squares linear regression, but the cost function can be altered as in the cases of the embedded methods above or weighted least squares linear regression.
Model validation can be carried out by calculating cross validated single parameter measures like R$^2$ scores which carry information about goodness of fit and overfitting.

\textbf{Importantly, the fit effective Hamiltonian carries with it a quantitative measure of its validity.}
The ability to quantify the accuracy of an effective Hamiltonian sets DMD apart from contemporary downfolding approaches which typically report only a final model.
Given the statistical nature of the fitting procedure, model quality assessment can be carried out and reported in exhaustive detail if one is interested.
Typically, however, simple single parameter measures like cross validated R$^2$ scores and root mean squared errors are used to quantify the accuracy of the effective Hamiltonian.

\subsection{Fixed-node diffusion Monte Carlo (FN-DMC)}
\textbf{Diffusion Monte Carlo (DMC) is a quantum Monte Carlo method which projects out the ground state of a real-space Hamiltonian given some initial trial wave function.}
Consider a trial wave function $|\Psi_T\rangle$ and the Hamiltonian $\hat{H}$ with ground state $|\Phi_0\rangle$. 
Applying the projector $e^{-\tau \hat{H}}$ as $\tau \rightarrow \infty$ to $|\Psi_T \rangle$
\begin{equation}
\lim_{\tau \rightarrow \infty} e^{-\tau \hat{H}} |\Psi_T\rangle 
\equiv \lim_{\tau \rightarrow \infty} |\Psi_\text{DMC}(\tau)\rangle \propto \langle \Phi_0|\Psi_T\rangle |\Phi_0\rangle,
\end{equation}
projects out the ground state as long as the trial wave function is not orthogonal to the ground state. 
DMC provides a stochastic implementation of this projector method and involves moving samples from the trial function $\Psi_T(R)$ using the Green function $G(R, R^\prime, \tau) = \langle R | e^{-\tau(\hat{H} - E_T)} | R^\prime \rangle$. 
Since $\hat{H} = \hat{T} + \hat{V}$, kinetic and potential energy terms, the Green function is approximated by a Trotter expansion $G(R, R^\prime, \tau) = \langle R | e^{-\tau(\hat{H} - E_T)} | R^\prime \rangle \sim \Big[e^{-d\tau(\frac{V(R) + V(R^\prime)}{2} - E_T)} \langle R| e^{-d\tau\hat{T}}|R^\prime \rangle + O(d\tau^2) \Big]^N $ where $d\tau = \tau/N$.
This expansion can be interpreted as an interative procedure where samples are moved N times with a small timestep $d\tau$ until convergence.
The constant $E_T$ is a trial energy used to control the normalization of $\Psi_\text{DMC}(\tau, R)$ and is updated at each move.

\textbf{DMC, however, suffers from a fermion sign problem which is alleviated via a fixed-node approximation.}
Under the fixed-node approximation the nodal surface of $\Psi_\text{DMC}(\tau, R)$ is forced to match that of the initial trial wave function for all $\tau$.
This approximation makes FN-DMC variational, and will only return the exact ground state of $\hat{H}$ if the nodal surfaces of $|\Psi_T\rangle$ and $|\Phi_0\rangle$ are identical.

\textbf{We take advantage of the variational nature of FN-DMC to sample the low-energy states necessary for DMD.}
Consider a set of trial wave functions with varying nodal structures.
The final projected $\Psi_\text{DMC}$ for these different trial functions will be the lowest energy states in $\mathcal{H}$ for the given nodal structures.
If the initial trial wave functions are appropriately chosen, the final projected states will be low-energy states within $\Psi_\text{DMC}$.
We calculate the expectation values $E[\Psi_\text{DMC}], \{d_k[\Psi_\text{DMC}]\}$ necessary for the model regression using a mixed estimator.
Details for calculating reduced density matrix elements in FN-DMC can be found in Wagner.

@@

\section{Proposed work}
My goal is to develop low-energy effective theories for the square lattice 1-band Hubbard model in order to map out the phases and transitions in the U-T space.
Recent attempts at solving this problem involve constructing single particle properties like densities of states using approximate numerical or effective theory techniques to classify different phases.
While instructive, reliance on single particle properties to identify phases instead of studying changes in the many-body Hilbert space and low-energy effective theory can lead to ambiguities regarding the nature of the phases.
I believe the DMD method can remove these ambiguities as the method explicitly involves studying changes in the low-energy Hilbert space and provides us with accurate many-body effective theories which can precisely differentiate different phases.

My first step would be to show whether the model supports more than one phase at T/t = 0 with respect to U, and if so what the features of these phases and the transition between them are.
I will investigate in particular the hypothesis that for $U/t<<1$ the model exhibits gapped Slater AFM behavior and for $U/t>>1$ Mott-Heisenberg behavior and whether the transition between them is a genuine phase transition or simply a crossover due to rearrangement of a few states.
To do so I will initially develop effective theories using DMD at two points U/t = 2, 12 and T/t = 0, allowing me to argue the existence or non-existence of multiple phases and investigate the properties of each phase.
I will follow by developing theories for various intermediate U in order to locate and study the transition between the two phases, if it exists, focusing on the hypothesis of a crossover U$^*/t\sim 4 - 5$ between the Slater and Mott-Heisenberg regimes.
Repeating this exercise for finite temperature I will build four more models at U/t = 2, 12 and T/t = 0.05, 0.20 to investigate the finite temperature phases at strong- and weak-coupling, in particular the claim of pseudogap and Fermi liquid phases at T/t = 0.05, 0.20 at weak coupling.
Filling in the U-T diagram will allow me to locate and study thermal transitions in the system. 
Building the T$>$0 effective theories will require special attention as they must satisfy the Mermin-Wagner theorem.

Similar to the case of the \textit{ab-initio} downfolding, I will need to construct low-energy spaces, sample the constructed spaces, build effective theories and respective bases, and fit candidate models.
I will begin from a starting point of highest control, namely for system sizes small enough where we can exactly diagonalize the Hubbard model.
In this case selecting $\mathcal{LE}$ can be done exactly as the span of some chosen $N$ lowest energy eigenstates. 
One can guarantee that the sample set fully encapsulates all low-energy degrees of freedom by sampling states parameterized as $|\Psi\rangle = \sum_{j=1}^N c_j |\Phi_j\rangle$ where $|\Phi_j\rangle$ are eigenstates of the Hubbard Hamiltonian.
Basis construction for the Hubbard Hamiltonian is also explicitly clear as any possible single particle basis we can use is just an orthogonal transformation of the basis the Hubbard model is defined on.
In order to include temperature into the mix one can work with the full model Hilbert space such that $\mathcal{LE} = \mathcal{H}$ but generate samples within the space using a Monte Carlo sampling method according to a Boltzmann probability distribution with variable parameter $\beta = 1/k_BT$, biasing the final model towards describing more accurately the energy functional on the states with highest weight.
Finite size extrapolation to the thermodynamic limit requires working with lattice sizes for which exact diagonalization is no longer feasible, and approximate methods like lattice variational Monte Carlo (VMC) will be employed to sample low-energy states as well as calculate energies and RDM elements.
In the weak and strong coupling regimes typical choices for varitional wave functions are non-interacting and Gutzwiller trial functions respectively, but other post-Gutzwiller and generalized Hartree Fock (GHF) trial functions have been shown to give accurate results in the moderate coupling regimes.
A rigorous prescription for finite size extrapolation specifically for the 1-band square lattice Hubbard model will be followed to extend our conclusions to the thermodynamic limit.

\begin{figure}[H]
\centering
\includegraphics[width=0.6\linewidth]{figs/eff_eig_prelim.pdf}
\caption{Results of weighted single parameter model fits to the eigenstates of a Hubbard model for various U and temperature. Darker hues indicate a higher R$^2$ score. Vertical dashed line denotes the bulk Slater to Mott-Heisenberg transition U$^*$/t = 4.7 calculated in [ref].}
\label{fig:Hubbard}
\end{figure}

I will conclude by showing some very preliminary results on model fitting for the half-filled, 2x2 square Hubbard model for U: 0 - 12 which put the outlined approach to model fitting with exact diagonalization at T$\geq 0$ into practice.
I started off by exactly diagonalizing each model, defining $\mathcal{LE} = \mathcal{H}$, and treating the full sample space as just the eigenstates of the model.
I then fit three single parameter models to the eigenstates: A) 
A "effective t" model $t_\text{eff} \sum_{\langle i, j \rangle, \sigma} c_{i\sigma}^\dagger c_{j\sigma} + h.c.$, B) An "effective U" model $U_\text{eff} \sum_{i} n_{i\uparrow} n_{i\downarrow}$, and C) An "effective J" model $J_\text{eff} \sum_{\langle i,j \rangle} \vec{S}_i \cdot \vec{S}_j$ using a weighted linear regression with weights $e^{-\beta E}$.
The basis for all of these operators are just the bare Hubbard Hamiltonian basis.
The weighted linear regression approximates the effect of generating samples using a Markov Chain Monte Carlo method with a Boltzmann distribution.
At low-temperatures there is a crossover between the weak coupling regime, best described by an effective t model, and the strong coupling regime, best described by an effective J model, around U$^*/t \sim 5$ providing some evidence of a Slater to Mott-Heisenberg transition.
In the crossover region T = 0 and U = 4 - 8 both models perform poorly and is a region of particular interest when developing our low-energy theories.
At high temperatures there is a crossover between the effective t and effective U models which can also be investigated.

>>>>

\section{Introduction}
Systematically developing model Hamiltonians which can accurately describe the energies and properties of the lowest lying eigenstates for strongly correlated electron systems is a pressing problem in modern condensed matter physics.

The necessity for non-perturbative, many-body models in describing the low-lying excited states of strongly correlated electron systems poses a difficult challenge when developing effective theories for real world systems.

Common approaches to building suitable model Hamiltonians for strongly correlated materials involve using single particle theories like density functional theory (DFT).

We propose using a DMD procedure to develop accurate many-body effective theories for single (SLG) and bi-layer graphene (BLG) systems as a step towards building a many-body model for twisted bilayer graphene (TBLG).

A core problem of condensed matter physics is understanding how to systematically develop low-energy effective models for strongly correlated electron systems. 
In the forefront of modern research are materials like high temperature superconductors and heavy fermion systems for which simple effective theories have not yet been found. 
For example, the cuprate family of high-Tc superconductors have an intricate phase diagram in the doping-temperature space with anti-ferromagnetic (AFM) insulating, superconducting, psuedo-gap, "bad metal", and Fermi-liquid like phases of which only the first and last are well understood.
The iron-pnictides also have superconducting domes, non-Fermi liquid like metallic phases, a spin-density-wave antiferommagnetic phase, all under various dopings and temperatures. 
Heavy-fermion materials have a purported quantum critical point between an AFM and Fermi-liquid phase which houses a superconducting dome, leading to another phase diagram for which building model Hamiltonians is very challenging.

@@

We propose using a density matrix downfolding (DMD) procedure to develop an accurate low-energy effective theory for twisted bilayer graphene (TBLG) beginning from the \textit{ab-initio} Hamiltonian.
TBLG exhibits strong correlation effects like unconventional superconductivity and many-body insulating states for extremely small, gate-tunable, dopings near charge neutrality for certain magic twist angles.
These strong correlation effects arise from the emergence of very flat bands near the Fermi level as one approaches the magic angles.
While some DFT calculations have generated effective single particle theories, no accurate many-body model Hamiltonian for TBLG which can describe these strong electronic correlations exists.
As discussed below, DMD provides us the unique ability to generate such accurate many-body models directly from the \textit{ab-initio} Hamiltonian.

\section{Methods}
\subsection{Density Matrix Downfolding (DMD)}
We begin by defining what we mean by a low-energy effective Hamiltonian and downfolding.
Consider a Hamiltonian $\hat{H}$ defined on a Hilbert space $\mathcal{H}$.
A low-energy effective Hamiltonian is an operator $\hat{H}_\text{eff}$ which can accurately reproduce the spectra, up to a constant, and eigenstates of $\hat{H}$ on a subpsace $\mathcal{LE}$ of the Hilbert space $\mathcal{H}$.
Here $\mathcal{LE}$ is taken to be the span of the N lowest energy eigenstates of $\hat{H}$, and $\hat{H}_\text{eff}$ is a linear combination of Hermitian operators $\hat{d}_k$ with a constant energy shift $E_0$
\begin{equation}
\hat{H}_\text{eff} = \sum_{k} g_k \hat{d}_k  + E_0.
\label{eq:Heff}
\end{equation}
The objective of Hamiltonian downfolding is to construct such an $\hat{H}_\text{eff}$ given $\hat{H}, \mathcal{H}$.

The key insight of DMD is that the Hamiltonian downfolding problem can be mapped onto a linear regression problem.
It has been shown that the above definition of $\hat{H}_\text{eff}$ is equivalent to the following: a low-energy effective Hamiltonian is an operator on a subspace $\mathcal{LE}$ of $\mathcal{H}$ such that 
\begin{equation}
\begin{split}
\forall |\Psi\rangle \in \mathcal{LE},\ (E[\Psi] \equiv \langle \Psi|\hat{H} | \Psi \rangle)  = \\ (E_\text{eff}[\Psi] \equiv \langle \Psi | \hat{H}_\text{eff} | \Psi \rangle) + \epsilon[\Psi]
\end{split}
\label{eq:DMD}
\end{equation}
where $\epsilon[\Psi]$ is the error in the effective theory.
Given the general linear form of $\hat{H}_\text{eff}$ seen in \eqref{eq:Heff}, 
the task of building an effective Hamiltonian then reduces to fitting a linear model that minimizes the error $\epsilon[\Psi]$ over $\mathcal{LE}$.
The independent variables (descriptors) in the fitting are expectation values of the operators $d_k[\Psi] \equiv \langle \Psi |\hat{d}_k|\Psi \rangle$ and the target variable is $E[\Psi]$.

This linear regression problem can be tackled in three steps, beginning with sampling states from $\mathcal{LE}$.
Ideally the linear regression would be conducted over the entire space $\mathcal{LE}$. 
In practice, the space is too large, and instead a representative sample set is drawn and used for fitting.
For each sampled state $|\Psi_s\rangle$, the quantities $E[\Psi_s], \{d_k[\Psi_s]\}$ are calculated to be used in the linear regression.
Optimal sampling schemes vary between systems and can significantly reduce the number of samples required to achieve robust statistics.

Generally, however, one cannot sample the true low-energy subspace and low-energy wave functions are sampled from an approximate subspace $\mathcal{LE}^\prime$.
A model fit to an arbitrary subspace of $\mathcal{H}$ will generally fail in describing the low-energy excitations of $\hat{H}$.
Rather, the choice of approximate subspace must be made carefully to ensure the fit model is transferable to the true low-energy space.
We describe in detail later how we selected $\mathcal{LE}^\prime$ for the CuO molecule and the scheme used to draw samples from it.

Next, a set of candidate descriptors which form the effective Hamiltonian are selected.
Descriptors will typically be selected if their values co-vary with the energy over the set of sampled states. 
Knowledge of the low-energy degrees of the freedom from experiment and physical intuition can help restrict the set of candidate descriptors further.
In principle there is no restriction on the form of the candidate operators other than Hermiticity, but second quantized operators are commonly used.

Typically, candidate operators are written in second quantized form, and an appropriate single particle basis to express these operators must be constructed.
An appropriate choice of single particle basis can simplify the representation of the effective theory and typically depends on the system of interest.
Common choices include Bloch waves and molecular orbitals for delocalized basis elements or Wannier and intrinsic atomic orbitals for localized ones.
The selected single particle basis can be validated by ensuring it accurately describes single particle properties among the sampled states like the total number of electrons in the system.

Finally, the sampled data are used to fit the coefficients $\{g_k\}$ by linear regression.
A typical fitting workflow involves feature selection, model regression and model validation.
Feature selection methods for choosing appropriate descriptors can range from wrapper methods like orthogonal matching pursuit and principal component analysis to embedded methods like LASSO, ridge and elastic net regularizations. 
Fitting the effective Hamiltonian usually involves an ordinary least squares linear regression, but the cost function can be altered as in the cases of the embedded methods above or weighted least squares linear regression.
Model validation can be carried out by calculating cross validated single parameter measures like R$^2$ scores which carry information about goodness of fit and overfitting.

Importantly, the fit effective Hamiltonian carries with it a quantitative measure of its validity.
The ability to quantify the accuracy of an effective Hamiltonian sets DMD apart from contemporary downfolding approaches which typically report only a final model.
Given the statistical nature of the fitting procedure, model quality assessment can be carried out and reported in exhaustive detail if one is interested.
Typically, however, simple single parameter measures like cross validated R$^2$ scores and root mean squared errors are used to quantify the accuracy of the effective Hamiltonian.

\subsection{Fixed-node diffusion Monte Carlo (FN-DMC)}
Diffusion Monte Carlo (DMC) is a quantum Monte Carlo method which projects out the ground state of a real-space Hamiltonian given some initial trial wave function.
Consider a trial wave function $|\Psi_T\rangle$ and the Hamiltonian $\hat{H}$ with ground state $|\Phi_0\rangle$. 
Applying the projector $e^{-\tau \hat{H}}$ as $\tau \rightarrow \infty$ to $|\Psi_T \rangle$
\begin{equation}
\lim_{\tau \rightarrow \infty} e^{-\tau \hat{H}} |\Psi_T\rangle 
\equiv \lim_{\tau \rightarrow \infty} |\Psi_\text{DMC}(\tau)\rangle \propto \langle \Phi_0|\Psi_T\rangle |\Phi_0\rangle,
\end{equation}
projects out the ground state as long as the trial wave function is not orthogonal to the ground state. 
DMC provides a stochastic implementation of this projector method and involves moving samples from the trial function $\Psi_T(R)$ using the Green function $G(R, R^\prime, \tau) = \langle R | e^{-\tau(\hat{H} - E_T)} | R^\prime \rangle$. 
Since $\hat{H} = \hat{T} + \hat{V}$, kinetic and potential energy terms, the Green function is approximated by a Trotter expansion $G(R, R^\prime, \tau) = \langle R | e^{-\tau(\hat{H} - E_T)} | R^\prime \rangle \sim \Big[e^{-d\tau(\frac{V(R) + V(R^\prime)}{2} - E_T)} \langle R| e^{-d\tau\hat{T}}|R^\prime \rangle + O(d\tau^2) \Big]^N $ where $d\tau = \tau/N$.
This expansion can be interpreted as an interative procedure where samples are moved N times with a small timestep $d\tau$ until convergence.
The constant $E_T$ is a trial energy used to control the normalization of $\Psi_\text{DMC}(\tau, R)$ and is updated at each move.

DMC, however, suffers from a fermion sign problem which is alleviated via a fixed-node approximation.
Under the fixed-node approximation the nodal surface of $\Psi_\text{DMC}(\tau, R)$ is forced to match that of the initial trial wave function for all $\tau$.
This approximation makes FN-DMC variational, and will only return the exact ground state of $\hat{H}$ if the nodal surfaces of $|\Psi_T\rangle$ and $|\Phi_0\rangle$ are identical.

We take advantage of the variational nature of FN-DMC to sample the low-energy states necessary for DMD.
Consider a set of trial wave functions with varying nodal structures.
The final projected $\Psi_\text{DMC}$ for these different trial functions will be the lowest energy states in $\mathcal{H}$ for the given nodal structures.
If the initial trial wave functions are appropriately chosen, the final projected states will be low-energy states within $\Psi_\text{DMC}$.
We calculate the expectation values $E[\Psi_\text{DMC}], \{d_k[\Psi_\text{DMC}]\}$ necessary for the model regression using a mixed estimator.
Details for calculating reduced density matrix elements in FN-DMC can be found in Wagner.

@@

\section{Proposed work}
My proposed plan involves building a sequence of model Hamiltonians for increasing complicated systems using DMD, working towards an accurate many-body model for TBLG.

The first system I will develop a model for is the benzene molecule.

While relatively simple, this molecule shares many similarities with and forms the basic unit of SLG.

As such, developing a model for benzene will help me become familiar with model fitting for 2-D carbon based systems.

The candidate descriptors I will use for DMD are motivated by a previous calculation on the benzene molecule.

A new constrained variational Monte Carlo (CVMC) method will be used to sample the low-energy wave functions necessary for DMD.

Note also that a model for benzene has already been developed using DMD.

This makes benzene a useful system not only for my personal learning, but as a benchmarking system for new methods like CVMC which have been developed recently.

The second system I will be interested in working on is SLG.

The model I am interested in working on will be an extension of a previous model developed for SLG using DMD.

I will extend the previous model by introducing long range density-density interactions into the effective theory.

The candidate model I anticipate to work with is an extended Hubbard model of the form.

CVMC will be used to sample the low-energy wave functions necessary to fit the model.

The last systems I will develop models for are AA and AB stacked BLG.

Being the simplest BLG configuration, these models will form a starting point for a more thorough development of models for TBLG.

I anticipate that the candidate descriptors should include interlayer couplings as well as the terms seen in SLG.

I will again use CVMC to generate the low-energy wave functions for the model regression.

The final four models can be used to investigate the transferrability of model parameters between these different carbon-based systems.

I will conclude by briefly discussing potential avenues of further study in using DMD to develop model Hamiltonians for TBLG.

My proposed plan involves building a sequence of model Hamiltonians for increasingly complicated systems using DMD, working towards a model for TBLG.
I will begin with the basic building block of SLG, the benzene molecule, as an exercise to become familiar with the structure and chemistry of graphene like materials.
The benzene molecule consists of a hexagonal carbon ring capped by hydrogen atoms. 
It shares many features with SLG such as delocalized $\pi$ electrons and long-range charge density-density interactions.
Alongside the fact that previous DMD work has already developed an accurate model for benzene, this molecule provides a useful playground for building my familiarity with the chemistry and effective interactions in graphene like systems.

The benzene model fitting will also serve as a benchmark study for new model fitting methods developed since the first DMD calculations on it.
I will be particularly interested in testing a constrained variation Monte Carlo (CVMC) sampling method which allows us to target low-energy states with specified descriptor values.
This sampling method will be the core sampling method used in all the proposed work, with the calculations on benzene serving as an familiarizing exercise for me.
The results of my model fit using CVMC can be validated by comparing the effective theory I have built to the many-body model previous developed for benzene.

The next step will be developing a model with long range interactions for SLG, expanding on previous work.
An on-site effective Hubbard model has already been built using DMD for single-layer graphene with parameters $U/t \sim 2.00$ eV.
Following the example of the benzene molecule, however, one would expect that longer range density-density interactions would be important in describing SLG more accurately.
The goal of this step of the project would be to develop a model of the following form:
\begin{equation}
\sum_{ij} t_{ij} c_i^\dagger c_j + h.c. +  \sum_i U n_{i\uparrow} n_{i\downarrow} + \sum_{ij} V_{ij} n_i^\dagger n_j
\end{equation}
with hopping terms $t_{ij}$, an on-site Hubbard $U$, and long range density-density interaction $V_{ij}$.

The last part of my plan involves developing a model for AA and AB stacked bilayer graphene.
These stackings are the simplest forms of bilayer graphene as the supercell is much smaller than those for even moderate commensurate twist angles.
The relative shift in distance among the atoms between the AA and AB stackings will serve as a first step to understand how the interlayer coupling between graphene layers may depend on the distance between atoms in the layers.
The final model to fit for these cases would be identical to the one above, but with additional interlayer hopping terms.

The end result of my proposed project is four model Hamiltonians for the  benzene molecule, SLG, AA stacked BLG, and AB stacked BLG.
These four models will include, at least, nearest neighbor hoppings, on-site Hubbard interactions, and long range density-density interactions.
For the AA and AB stacked BLG we will also consider interlayer couplings starting with just hoppings between the layers.
Each model will come with fit parameter values, an appropriate single particle basis, and a measure of the goodness of fit for the model.
These models can be solved and compared to experiment.

With these models we can resolve an interesting question: are model parameters transferable between benzene, SLG and BLG?
The question of transferability of model parameters between SLG and BLG is quite interesting because it may inform us about how much the behavior of each graphene sheet changes once we stack them.
If the interlayer coupling does not affect the effective behavior on the single sheets, then one can easily treat the final model for bi-layer graphene in two independent pieces: an accurate model for the graphene sheets with a position dependent interlaying coupling.
However, if the interlayer coupling plays a large role in the effective theory then it may be impossible to separate the BLG model into these two pieces.
A similar study can be done for the single particle basis, giving some information about orbital relaxation going from SLG to BLG.

My work will provide a useful baseline to begin studies on TBLG.
The approach to TBLG is still very far away and will require all of the previous steps above. 
Afterwards, work on TBLG will require building effective theories of BLG with multiple twist angles, or at least varying local distances between atoms.
The goal here is to develop models for BLG where the parameters, like interlayer coupling, are defined continuously as a function of atomic distance.
The resulting interacting lattice model can then be solved over very large supercell sizes and a second step of downfolding can be done.
The goal of this second step would be to downfold the lattice model into a continuum model over the length scale of the Moire patterns seen in small twist angle BLG.
Work on TBLG would be highly exploratory, and the steps detailed here are very speculative, outlining a possible avenue of work continuing after my thesis.
